---
output: md_document
---


#### ⠀
#### ⠀

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(dplyr)
library(olsrr)
library(corrplot)
library(ggplot2)
library(MASS)
library(imager)
library(magick)

# Title page
titlepage = image_read('images/titlePage.jpg')
image_resize(titlepage,"1800x")

# Just a note that there's a lot of "####" in this file, they have space character after them to create gap on the output.
```

#### ⠀
#### ⠀


### Abstract

Fishes are among the most distinctive group of vertebrates consisting of more than 33,000 known species and they are salient in various sectors such as farming, processing, products, and marketing. Further, a vital part after harvesting fish is to ascertain their proportions and weight, where measuring individual weights can be expensive and time-consuming. 

Therefore, the fishing industry necessitates methods to identify the weight of different species to operate proficiently. This project aims to explore statistical multilinear regression modeling to determine ways to identify the fish weight by its dimensions for commonly seen fish species at a fish market. 

#### ⠀
#### ⠀

### Introduction

The fishing industry plays an important role directly in the livelihood of millions, it was reported that at the start of the 21st century an estimated 250 million people worldwide were directly employed by the industry and an estimated one billion people rely on fish as their primary source of protein. There has been radical advancement of fishing technology throughout history which includes larger shipping vessels, sophisticated fishing equipment, numerous food preservation methods etc. However, these technologies are costly and heavily time-consuming and thus, this project investigates cost-effective methods to weigh fishes where the fish species and their proportions are known. The resulting models immensely benefit the entities that need to weigh individual fish on a day-to-day basis.

The fish dataset is composed of 159 observations and one additional observation is introduced to the dataset that impacts the analysis which will be discussed in depth in the sections that follow. 
The following are the questions of interest: 

1.	Is there a linear relationship between the Lengths variables, Height, Width, and Weight for all species? 

2.	If so, does this relationship vary by species? 

3.	Is there an effective model that can predict the individual fish weights for a specified species by using the fish dimensions? 


#### ⠀
#### ⠀



### Data description
#### ⠀

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Prints fish images that explains the dimensions side by side
image2 = image_read('images/fish1.jpg')
image_resize(image2,"700x")
image3 = image_read('images/fish2.jpg')
image_resize(image3,"700x")
```


```{r echo=FALSE}
# Description table for all the variables
image4 = image_read('images/table.jpg')
image_resize(image4,"1300x")
```


```{r message=FALSE, warning=FALSE, include=FALSE}
## Reading the fishdata and renaming the first column header to species
FishMarket <- read.csv(file = "350Data.csv", header = T)
names(FishMarket)[names(FishMarket)=="ï..Species"] <- "Species"
```

#### ⠀
#### ⠀
#### First impressions with various plots and the new data observation:
#### ⠀


```{r include=FALSE}
newObsSpecies <- 'Whitefish'
newObsWeight <- median(FishMarket$Weight[56:61])
newObsLength1 <- median(FishMarket$Length1[56:61])
newObsLength2 <- median(FishMarket$Length2[56:61])
newObsLength3 <- median(FishMarket$Length3[56:61])
newObsHeight <- median(FishMarket$Height[56:61])
newObsWidth <- median(FishMarket$Width[56:61])

newObservation <-  list(newObsSpecies, newObsWeight, newObsLength1, newObsLength2, newObsLength3, newObsHeight, newObsWidth)
FishMarket <- rbind(FishMarket, newObservation)

```




```{r echo=FALSE}
# Histogram
ggplot(data = FishMarket) + geom_bar(aes(Species))
```


The aforementioned new additional data point is created for the Whitefish species primarily due to its low count seen in the histogram above. Furthermore, the rest of its numerical variables are calculated by the median average and then appended to the fish dataset. 


#### ⠀


```{r echo=FALSE}
# Exploratory Data Analysis

# Boxplot

ggplot(data = FishMarket,aes(Species,Weight,fill = Species))+
  geom_boxplot(position = "dodge2")

# Checking for linear association between weight and the dimensions of the fish

ggplot(data = FishMarket,mapping = aes(x=Length1,y=Weight,color=Species))+
  geom_point()
ggplot(data = FishMarket,mapping = aes(x=Length2,y=Weight,color=Species))+
  geom_point()
ggplot(data = FishMarket,mapping = aes(x=Length3,y=Weight,color=Species))+
  geom_point()
ggplot(data = FishMarket,mapping = aes(x=Width,y=Weight,color=Species))+
  geom_point()
ggplot(data = FishMarket,mapping = aes(x=Height,y=Weight,color=Species))+
  geom_point()

```



From the plots, it is visible that weight (response variable) and dimensions of the fish (predictor variables) are positively 
related, however, the association is not linear. The proper type of transformation will be implemented at a later stage.

#### ⠀
#### ⠀

### Methods 

```{r echo=TRUE}
# Checking for the presence of multicollinearity in the independent variables

# Correlation between variables
correlation <- cor(FishMarket[,-1])
corrplot(correlation,method = "number")

# Eigenvalues
X <- cbind(FishMarket$Length1,FishMarket$Length2,FishMarket$Length3,
           FishMarket$Height,FishMarket$Width)
XX=t(X)%*%X
lambda = eigen(XX)$values
indices=max(lambda)/lambda
names(FishMarket[,which(indices>1000)])
```


The correlation plot highlights the high correlation of 0.99 between all length variables. Also, applying the eigenvalues method for all the variables, the output shows that the indices values are greater than 1000 for all of the lengths which means multicollinearity will be an issue if all of the lengths are included. Thus, only Length3 will be considered for the multilinear regression model because it is the total length of the fish. 

#### ⠀
#### ⠀

```{r}
# Three different models are compared to check which dimension variable is the most significant:
test.model <- lm(Weight ~ Length3+Width+Height - 1,data = FishMarket)

dummy.lg.model <- lm(Weight ~ Length3+Width+Height+Length3:Species - 1,data = FishMarket)
dummy.wd.model <- lm(Weight ~ Length3+Width+Height+Width:Species - 1,data = FishMarket)
dummy.hg.model <- lm(Weight ~ Length3+Width+Height+Height:Species - 1,data = FishMarket)

print(anova(test.model,dummy.lg.model))
print(anova(test.model,dummy.wd.model))
print(anova(test.model,dummy.hg.model))
```
#### ⠀

Hypothesis for all the ANOVA tests: 

Null hypothesis: The interaction terms are equal to 0 

Alternate hypothesis: The interaction terms are non-zero 

The ANOVA tests are conducted to compares the different models that include the species interaction term and models without the interaction term. All the p-values are less than alpha of 0.05 and we reject the null hypothesis that the interaction terms are zero and conclude that there is a relationship between the dimensions of the fish and the species interaction terms. Note that the Length:Species is the most significant interaction term because it has the smallest p-value.

#### ⠀
#### ⠀


```{r}
## Species correlation plot
Dummy.data <- fastDummies::dummy_cols(FishMarket, select_columns = "Species")
Dummy.data <- Dummy.data[,-c(1,3,4,5,6,7)]
corrplot(cor(Dummy.data),method = "number")

```



The correlation plot of weight VS. the indicator species variables. From the plot, it is shown that there won't be any multicollinearity issues. 

#### ⠀
#### ⠀


```{r}
# Transformation Model:

# Original model
linear.model <- lm(Weight ~Length3+Height+Width,data = FishMarket)
summary(linear.model)

# Square root transformation on weight(response variable)
sq.trans.FishMarket <- FishMarket
sq.trans.FishMarket$Weight <- sqrt(sq.trans.FishMarket$Weight)
sq.trans.model <- lm(Weight ~Length3+Height+Width,data = sq.trans.FishMarket[,-1])
summary(sq.trans.model)

## Log transformation on weight
log.trans.FishMarket <- FishMarket
log.trans.FishMarket$Weight <- log(log.trans.FishMarket$Weight)
log.trans.FishMarket[which(log.trans.FishMarket$Weight==0),]
log.trans.FishMarket<-log.trans.FishMarket[-41,]
log.trans.model <- lm(Weight ~Length3+Height+Width,data = log.trans.FishMarket[,-1])
summary(log.trans.model)

```



Three different models are considered: no transformation, square root transformation on the weight variable and log transformation on the weight variable. From the output, we can conclude let the square root transformation is the best choice for the highest R-square value of 0.9709.

#### ⠀
#### ⠀


```{r}
# Variable Selection:
temp.model <- lm(Weight~Height + Width + Length3 + factor(Species) + 
                   Height:Species + Width:Species + Length3:Species,data = FishMarket)

# Stepwise regression model (AIC)
stepwise <- stepAIC(temp.model, direction = "both", trace = FALSE)
summary(stepwise)

```



The stepwide model selection by Akaike Information Criterion (AIC) is used and the best model is with Height, Width3, Length3, Species and the Length3:Species interaction term. 

#### ⠀
#### ⠀



```{r}
# Cross Validation：

# Data splitting
set.seed(350)
nsamp=ceiling(0.8*length(FishMarket$Weight))
training_samps=sample(c(1:length(FishMarket$Weight)),nsamp)
training_samps=sort(training_samps)
train_data <- FishMarket[training_samps, ]
test_data <- FishMarket[-training_samps, ]

train_data$Weight <- sqrt(train_data$Weight)
test_data$Weight <- sqrt(test_data$Weight)

# Fitting model
myModel <- lm(Weight~Height + Width + Length3 + factor(Species) + Length3:Species,data = train_data)
summary(myModel)

# Testing on the test data
preds <- predict(myModel,test_data)
plot(test_data$Weight,preds,main = "Test Data vs Prediction",xlab = "test data",ylab = "prediction")
abline(c(0,1),col="red")

```



Using the cross validation method, we set 80% of the data to be training samples which are used for building a model and the rest of the data is used for testing the accuracy of the model. The Test Data vs Prediction plot shows that the model is predicting effectively. 

#### ⠀
#### ⠀



```{r}
# Regression Analysis
plot(myModel,which = 1)


plot(myModel,which = 2)


plot(myModel,which = 3)


plot(myModel,which = 5)



```



The Residuals vs Fitted plot has most of the fitted residuals around zero. 

The linear assumption is met after the square root transformation on the response variable. 

The Normal Q-Q plot shows that the normal distributed assumption is met. 

The Standardized Residual plot shows that the constant variance assumption is met. 

The Residuals vs Leverage plot shows that the uncorrelated errors assumption is met and most of the data is within the dashed lines of Cook's distance. 

However, observations 14, 41 and 143 need further investigation because they have large residual and leverage value.

#### ⠀
#### ⠀


```{r}
# Residual Analysis 

# Cook distance
cook.dis<- cooks.distance(myModel)
head(cook.dis)
train_data [which(cook.dis>1),] #points with high leverage

# Standardized Residual
rstandard(myModel)
stdRsidual<-rstandard(myModel)
train_data[which(abs(stdRsidual)>3),]


```


After calculating the Cook's distance of each suspected observation, the output shows that none of the points have Cook's distance larger than 1, which means there is no high leverage point. However, observations 41 and 14 have standardized residuals that are larger than 3 which means they might be considered as outliers.


#### ⠀
#### ⠀

```{r}
# Remove Outlier
new_train_data <- train_data[-c(41,14),]

# Fitting model
new_myModel <- lm(Weight~Height + Width + Length3 + factor(Species) + Length3:Species,data = new_train_data)
summary(new_myModel)
summary(myModel)

# Testing on test data
preds <- predict(myModel,test_data)
plot(test_data$Weight,preds,main = "Test Data vs Prediction",xlab = "test data",ylab = "prediction")
abline(c(0,1),col="red")

```



The suspected outliers (observations 41 and 14) are removed from the fish dataset and a new model is fitted. The output shows that the coefficients and R-square of the new model do not alter by much so the observations 41 and 14 are kept in the dataset.

#### ⠀
#### ⠀

```{r}
# Robust Regression:
rb.model <- rlm(Weight~Height + Width + Length3 + factor(Species) + Length3:Species,data = train_data)
summary(rb.model)


# Points with least weight:
head(order(rb.model$w))

# Testing on test data
par(mfrow=c(1,2))
preds <- predict(myModel,test_data)
plot(test_data$Weight,preds,main = "Original Regression Model",xlab = "test data",ylab = "prediction")
abline(c(0,1),col="red")
preds.rb <- predict(rb.model,test_data)
plot(test_data$Weight,preds.rb,main = "Robust Regression Model",xlab = "test data",ylab = "prediction")
abline(c(0,1),col="blue")


```



Robust regression is applied on the train data, however, the weight is not less on the potential outliers 41 and 14. It could be because they are not influential to the original model. Comparing the Test Data vs Prediction plot of the original model and robust model, there is little difference and thus the original model is kept. 

#### ⠀
#### ⠀




#### Results

From the methods conducted in the last section, we can deduce the multilinear regression model for all the species: 

```{r echo=FALSE}
# The multilinear models for all the species
image4 = image_read('images/results.jpg')
image_resize(image4,"1200x")
```


For all species, the B0 is a constant and for every increment of one centimeter of the fish dimensions, the height increases by b1, the width increases by b2 and the length3 increases by b3. 

The results can be implemented by a user to determine the Weight of any of the 7 species by plugging in the dimensions of the fish in the model.


#### ⠀
#### ⠀

### Conclusion


Now in this final section, all the dots can be connected. The purpose of this project is to create a multilinear regression model that allows the user to calculate the individual weight of a fish while knowing the height, the width, the total length (length3) and the species type. Various statistical methods were implemented on the fish dataset to build adequate multilinear regression models for all species. 

These methods diligently answer our questions of purpose: 

1.	Is there a linear relationship between the Lengths variables, Height, Width, and Weight? 

    Yes, there is a linear positive relationship between the Length variables, Height, Width and Weight where square root transformation was applied to the Weight variable. 

2.	If so, does this relationship vary by species? 
    
    Yes, it varies for different species which is why the model has different intercepts and slopes based on the specific species.

3.	Is there an effective model that can predict the individual fish weights for a specified species by using the fish dimensions? 

    Yes, the 7 multilinear regression models in the result section are effective models that can predict the individual fish weights for a specified species by using the fish dimensions.


Thus, these 7 multilinear regression models are ready to be used by the fishing industry, particularly the entities who need to find the weight for Bream, Roach, Whitefish, Parkki, Perch, Pike and Smelt while knowing their dimensions. 

#### ⠀
#### ⠀



```{r echo=FALSE, message=FALSE, warning=FALSE}
image = image_read('images/findingweight.jpg')
image_resize(image,"1000x")
```



#### ⠀
#### ⠀



### Appendix

The dataset is extracted from: https://www.kaggle.com/aungpyaeap/fish-market

The R Markdown file and all additional files are available in the repository. 


#### ⠀
#### ⠀
#### ⠀
#### ⠀















